Trajectories

### 2022-09-07 docs/index.qmd
Report to Quarto

---
title: "Trajectories"
authors:
    - Gonzalo Garcia-Castro^[Universitat Pompeu Fabra, gonzalo.garciadecastro@upf.edu]
    - Daniela Avila-Varela^[Universitat Pompeu Fabra]
    - Ignacio Castillejo^[Universidad Autónoma de Madrid]
    - Núria Sebastian-Galles^[Universitat Pompeu Fabra]
output:
    html_document:
        code_folding: hide
        number_sections: true
        toc: true
    github_document:
        pandoc_args: --webtex
        number_sections: true
        toc: true
bibliography: "../references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    message = TRUE,
    echo = TRUE,
    warning = TRUE,
    include = TRUE,
    out.width = "80%",
    dpi = 250
)
options(
    knitr.kable.NA = "-",
    knitr.duplicate.label = "allow",
    ggplot2.discrete.fill = ggsci::pal_d3()(4),
    ggplot2.discrete.colour = ggsci::pal_d3()(4),
    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
    ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

```{r prepare, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# set params and load target contents
tar_load_globals()
tar_load_all()
theme_set(theme_custom())

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

@@

```{r items, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
pool <- multilex::pool %>% 
    drop_na(cognate)
n_categories <- distinct(pool, category) %>% nrow()
n_items <- length(unique(pool$item))
n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
n_item_cognate <- count(pool, language, cognate) %>% 
    pull(n) %>% 
    set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
```

@@

# Items {.tabset .tab-pills}

```{r items_frequency, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.cap="Lexical frequencies. Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."}

items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))

items %>%
    unnest(list) %>% 
    group_by(language, list) %>% 
    summarise_at(
        vars(freq, n_phon, lv), 
        lst(n = ~sum(!is.na(.)), mean, sd, min, max)
    ) %>% 
    select(-c(n_phon_n, lv_n)) %>% 
    rename(n = freq_n) %>% 
    gt(groupname_col = "language") %>% 
    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
    fmt_percent(matches("lv_")) %>% 
    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
    tab_spanner("Levenshtein", columns = starts_with("lv_")) %>% 
    cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
    cols_merge_range(col_begin = "n_phon_min", col_end = "n_phon_max") %>% 
    cols_merge_range(col_begin = "lv_min", col_end = "lv_max") %>% 
    cols_label(
        list = "",
        n = md("*N*"),
        freq_mean = md("Mean"),
        freq_sd = md("*SD*"),
        freq_min = md("Range"),
        n_phon_mean = md("Mean"),
        n_phon_sd = md("*SD*"),
        n_phon_min = md("Range"),
        lv_mean = md("Mean"),
        lv_sd = md("*SD*"),
        lv_min = md("Range")
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(everything())
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(everything())
    )

@@

```{r items-frequency, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}

items %>% 
    ggplot() +
    aes(
        x = freq,
        fill = language,
        colour = language,
        shape = language
    ) +
    stat_slab(
        slab_colour = "white",
        slab_size = 1,
        alpha = 0.5,
    ) +
    stat_pointinterval(
        aes(y = -0.1),
        position = position_dodge(width = 0.1),
        point_size = 3
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "Frequency (Zipf)",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent,  limits = c(-0.15, 1), breaks = seq(0, 1, 0.25)) +
    theme(
        legend.position = "top",
        legend.title = element_blank()
    )

@@

```{r items-nphon, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}

items %>%
    count(language, n_phon) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(
        x = n_phon,
        y = prop,
        fill = language,
        colour = language,
        shape = language
    ) +
    geom_col(
        position = position_dodge(width = 0.95)
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "# Phonemes",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
    scale_x_continuous(breaks = seq(0, 20, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(),
        legend.title = element_blank()
    )

@@

```{r items-lv, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}

items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv_cut = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv_cut) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(x = lv_cut, y = prop) +
    geom_col(fill = ggsci::pal_d3()(4)[1]) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
    labs(
        x = "Levenshtein",
        y = "Probability density",
    ) +
    theme(
        legend.position = "top"
    )

@@

```{r items-frequency-phonemes, message=FALSE, warning=FALSE, paged.print=FALSE}
items %>% 
    drop_na(freq) %>% 
    ggplot(aes(n_phon, freq, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r items-frequency-levenshtein, message=FALSE, warning=FALSE, paged.print=FALSE}
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(freq, lv, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.5) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r items-phonemes-lv, message=FALSE, warning=FALSE, paged.print=FALSE}
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(n_phon, lv, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Levenshtein similarity", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

## {-}

# Participants {.tabset .tab-pills}

```{r participants_time, message=FALSE, warning=FALSE, paged.print=FALSE}
participants <- left_join(
    participants, 
    select(multilex_data$logs, id, time, edu_parent, dominant_language = dominance)
)

participants %>% 
    group_by(time_stamp, dominant_language) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    group_by(dominant_language) %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n, colour = dominant_language, fill = dominant_language)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 1) +
    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
    scale_y_continuous(breaks = seq(0, 400, 50), limits = c(0, 400)) +
    guides(colour = guide_legend(ncol = 2)) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
    )

@@

```{r participants-age, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = pal_d3()(3)[3]) +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(
        x = "Age (months)",
        y = "# participants"
    ) +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()

    )
```

@@

```{r participants-lp, message=FALSE, warning=FALSE, paged.print=FALSE}

participants %>% 
    mutate(
        age = round(age),
        id = paste(id, "(", time, ")"),
        rank = rank(doe_catalan)
    ) %>% 
    pivot_longer(
        starts_with("doe_"), 
        names_to = "language",
        values_to = "doe",
        names_prefix = "doe_",
        names_transform = list(language = str_to_sentence)
    ) %>% 
    ggplot() +
    aes(x = doe, y = reorder(id, -rank), fill = language) +
    geom_col(width = 1, position = position_fill(), colour = "white", size = 0.1) +
    geom_vline(xintercept = seq(0, 1, 0.25), colour = "grey") +
    labs(
        x = "Degree of Exposure (DoE)",
        y = "Participant",
        colour = "Language",
        fill = "Language"
    ) +
    scale_x_continuous(labels = percent) +
    theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.title = element_blank()
    )

@@

```{r participants_edu, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>%
    mutate(
        edu_parent = fct_explicit_na(as.factor(edu_parent)),
        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
    ) %>% 
    count(doe_catalan, edu_parent) %>% 
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(multilex_data$logs$edu_parent)
        )
    ) %>% 
    rename(total = n) %>% 
    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
    rename(Total = total) %>% 
    select(-"NA") %>% 
    gt() %>% 
    cols_label(
        doe_catalan = "DoE Catalan"
    ) %>% 
    summary_rows(
        columns = 2:7, 
        fns = list(Total = ~sum(., na.rm = TRUE)),
        formatter = fmt_number,
        decimals = 0
    ) %>% 
    fmt_missing(columns = 3:8, missing_text = "--") %>% 
    tab_spanner("Educational attainment", 3:8) %>% 
    tab_style(
        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
        locations = cells_body(columns = 9, rows = 1:10)
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

## {-}

# Data analysis {.tabset .tab-pills}

@@

$$\begin{aligned}
g[\mathrm{Pr}(y_{ij} \le a_s)] = \tau_s - \nu_{ij}
\end{aligned}$$

@@

$$
\begin{aligned}

@@

\end{aligned}
$$

@@

$$\begin{aligned}
\beta_{0} & \sim \mathrm{Normal}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
\beta_{1} &\sim \mathrm{Normal}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
\beta_{2-9} &\sim \mathrm{Normal}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
\theta_{0-9} & \sim \mathrm{Normal}(1, 0.1) & [\mbox{Group-level coefficient variability}]
\end{aligned}
$$

@@

## R code (`brms`) {.tabset}

@@

### {-}

## Stan code {.tabset}

Stan code generated by `brms::stancode`:

@@

### {-}

## {-}

# Results {.tabset .tab-pills}

@@

```{r raw, message=FALSE, warning=FALSE, paged.print=FALSE}
df %>%
    mutate(
        lv = cut_interval(lv, 5),
        age = floor(age)
    ) %>% 
    count(lv, age, response) %>%
    mutate(prop = n/sum(.$n)) %>% 
    ggplot(aes(age, prop, colour = response, fill = response)) +
    facet_wrap(~lv, nrow = 1) +
    geom_col(position = position_fill()) +
    geom_hline(yintercept = 0.5, colour = "black", size = 0.5) +
    labs(
        x = "Age (months)",
        y = "% responses", 
        colour = "Response",
        fill = "Response"
    ) +
    scale_fill_d3() +
    scale_colour_d3() +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
    theme(
        legend.position = "top",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

@@

```{r loos, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, eval=FALSE}
model_loos %>% 
    as_tibble() %>% 
    rownames_to_column("model") %>% 
    relocate(
        model, 
        matches("elpd_loo"),
        matches("p_loo"), 
        matches("looic"), 
        matches("diff")
    ) %>%
    mutate_all(as.numeric) %>% 
    gt() %>% 
    tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
    tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
    tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
    tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
    fmt_number(3:13) %>% 
    cols_label(
        model = "Model",
        # elpd
        elpd_loo = md("*ELPD*"),
        se_elpd_loo = md("*SE*"),
        subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
        # p
        p_loo = md("*p*"),
        se_p_loo = md("SE"),
        subsampling_se_p_loo = md("SE<sub>sub</sub>"),
        # looic
        looic = md("*LOO-IC*"),
        se_looic = md("*SE*"),
        subsampling_se_looic = md("SE<sub>sub</sub>"),
        # diff
        elpd_diff = md("*diff*"),
        se_diff = md("*SE*"),
        subsampling_se_diff = md("*SE<sub>sub</sub>*"),
    ) 
```

@@

## Random effects {.tabset .tab-pils}

@@

### {-}

@@

```{r marginal, message=FALSE, warning=FALSE, paged.print=FALSE}
nd <- expand.grid(
    n_phon_std = 0,
    freq_std = 0,
    age_std = seq(-4, 4, length.out = 30),
    doe_std = range(df$doe_std),
    lv_std = c(range(df$lv_std), 0)
)

@@

## {-}

>>>>

---
title: "Trajectories"
subtitle: "Lab notes"
date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
format:
    html:
        abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies. Floccia et al. (2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch). The mechanisms underlying this effect remain unknown. We explore two compatible scenarios. First, we test whether cognates are learnt earlier than non-cognates. This would account for the difference in vocabulary size associated to the amount of shared cognates across languages. Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap. This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
        toc: true
        smooth-scroll: true
        link-external-newwindow: true
        number-sections: true
        number-offset: 0
        code-fold: true
        code-overflow: scroll
        code-line-numbers: true
        code-copy: hover
        fig-dpi: 500
        reference-location: margin
        execute-dir: "C:/Users/U155880/Documents/trajectories/"
bibliography: "references.bib"
csl: "apa6.csl"
---

```{r}
#| label: setup
#| include: false
options(
    knitr.kable.NA = "-",
    knitr.duplicate.label = "allow",
    ggplot2.discrete.fill = ggsci::pal_d3()(4),
    ggplot2.discrete.colour = ggsci::pal_d3()(4),
    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
    ggplot2.continuous.colour = ggplot2::scale_color_gradient
)
```

```{r}
#| label: prepare
#| echo: true
#| message: false
#| warning: false
# set params and load target contents
targets::tar_load_globals()
targets::tar_load_everything()

# set custom ggplot2 theme
theme_set(theme_custom())

# resolve namespace conflicts
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

@@

```{r}
#| label: questionnaire
#| echo: true
#| message: false
#| warning: false
pool <- multilex::pool %>% 
    drop_na(cognate)
n_categories <- distinct(pool, category) %>% nrow()
n_items <- length(unique(pool$item))
n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
n_item_cognate <- count(pool, language, cognate) %>% 
    pull(n) %>% 
    set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))
```

@@

# Items

```{r}
#| label: items-summary
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Lexical frequencies. Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))

items %>%
    unnest(list) %>% 
    group_by(language, list) %>% 
    summarise_at(
        vars(freq, n_phon, lv), 
        lst(n = ~sum(!is.na(.)), mean, sd, min, max)
    ) %>% 
    select(-c(n_phon_n, lv_n)) %>% 
    rename(n = freq_n) %>% 
    gt(groupname_col = "language") %>% 
    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
    fmt_percent(matches("lv_")) %>% 
    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
    tab_spanner("# Phonemes", columns = starts_with("n_phon_")) %>% 
    tab_spanner("Levenshtein", columns = starts_with("lv_")) %>% 
    cols_merge_range(col_begin = "freq_min", col_end = "freq_max") %>% 
    cols_merge_range(col_begin = "n_phon_min", col_end = "n_phon_max") %>% 
    cols_merge_range(col_begin = "lv_min", col_end = "lv_max") %>% 
    cols_label(
        list = "",
        n = md("*N*"),
        freq_mean = md("Mean"),
        freq_sd = md("*SD*"),
        freq_min = md("Range"),
        n_phon_mean = md("Mean"),
        n_phon_sd = md("*SD*"),
        n_phon_min = md("Range"),
        lv_mean = md("Mean"),
        lv_sd = md("*SD*"),
        lv_min = md("Range")
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(everything())
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(everything())
    )
```

@@

```{r}
#| label: items-frequency
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>% 
    ggplot() +
    aes(
        x = freq,
        fill = language,
        colour = language,
        shape = language
    ) +
    stat_slab(
        slab_colour = "white",
        slab_size = 1,
        alpha = 0.5,
    ) +
    stat_pointinterval(
        aes(y = -0.1),
        position = position_dodge(width = 0.1),
        point_size = 3
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "Frequency (Zipf)",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent,  limits = c(-0.15, 1), breaks = seq(0, 1, 0.25)) +
    theme(
        legend.position = "top",
        legend.title = element_blank()
    )
```

@@

```{r}
#| label: items-nphon
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>%
    count(language, n_phon) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(
        x = n_phon,
        y = prop,
        fill = language,
        colour = language,
        shape = language
    ) +
    geom_col(
        position = position_dodge(width = 0.95)
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "# Phonemes",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
    scale_x_continuous(breaks = seq(0, 20, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(),
        legend.title = element_blank()
    )
```

@@

```{r}
#| label: items-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv_cut = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv_cut) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(x = lv_cut, y = prop) +
    geom_col(fill = ggsci::pal_d3()(4)[1]) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
    labs(
        x = "Levenshtein",
        y = "Probability density",
    ) +
    theme(
        legend.position = "top"
    )
```

@@

```{r}
#| label: items-frequency-phonemes
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>% 
    drop_na(freq) %>% 
    ggplot(aes(n_phon, freq, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: items-frequency-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(freq, lv, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.5) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: items-nphon-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(n_phon, lv, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Levenshtein similarity", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

# Participants

```{r}
#| label: participants-time
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
participants <- left_join(
    participants, 
    select(multilex_data$logs, id, time, edu_parent, dominant_language = dominance)
)

participants %>% 
    group_by(time_stamp, dominant_language) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    group_by(dominant_language) %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n, colour = dominant_language, fill = dominant_language)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 1) +
    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
    scale_y_continuous(breaks = seq(0, 400, 50), limits = c(0, 400)) +
    guides(colour = guide_legend(ncol = 2)) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
    )
```

@@

```{r}
#| label: participants-age
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = pal_d3()(3)[1]) +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(
        x = "Age (months)",
        y = "# participants"
    ) +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

@@

```{r}
#| label: participants-lp
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
participants %>% 
    mutate(
        age = round(age),
        id = paste(id, "(", time, ")"),
        rank = rank(doe_catalan)
    ) %>% 
    pivot_longer(
        starts_with("doe_"), 
        names_to = "language",
        values_to = "doe",
        names_prefix = "doe_",
        names_transform = list(language = str_to_sentence)
    ) %>% 
    ggplot() +
    aes(x = doe, y = reorder(id, -rank), fill = language) +
    geom_col(width = 1, position = position_fill(), colour = "white", size = 0.1) +
    geom_vline(xintercept = seq(0, 1, 0.25), colour = "grey") +
    labs(
        x = "Degree of Exposure (DoE)",
        y = "Participant",
        colour = "Language",
        fill = "Language"
    ) +
    scale_x_continuous(labels = percent) +
    theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.title = element_blank()
    )
```

@@

```{r}
#| label: participants-edu
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
participants %>%
    mutate(
        edu_parent = fct_explicit_na(as.factor(edu_parent)),
        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
    ) %>% 
    count(doe_catalan, edu_parent) %>% 
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(multilex_data$logs$edu_parent)
        )
    ) %>% 
    rename(total = n) %>% 
    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
    rename(Total = total) %>% 
    select(-"NA") %>% 
    gt() %>% 
    cols_label(
        doe_catalan = "DoE Catalan"
    ) %>% 
    summary_rows(
        columns = 2:7, 
        fns = list(Total = ~sum(., na.rm = TRUE)),
        formatter = fmt_number,
        decimals = 0
    ) %>% 
    fmt_missing(columns = 3:8, missing_text = "--") %>% 
    tab_spanner("Educational attainment", 3:8) %>% 
    tab_style(
        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
        locations = cells_body(columns = 9, rows = 1:10)
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

# Data analysis

@@

:::{.column-body-outset}
$$\begin{aligned}
g[\mathrm{Pr}(y_{ij} \le a_s)] = \tau_s - \nu_{ij}
\end{aligned}$$
:::

@@

:::{.column-body-outset}
$$
\begin{aligned}

@@

\end{aligned}
$$
:::

@@

:::{.column-body-outset}
$$\begin{aligned}
\beta_{0} & \sim \mathrm{Normal}(-0.25, 0.1) & [\mbox{Intercept/response category threshold}] \\
\beta_{1} &\sim \mathrm{Normal}(1, 0.1) & [\mbox{Age population-level coefficient}]\\
\beta_{2-9} &\sim \mathrm{Normal}(0, 1) & [\mbox{Rest of population-level coefficients}] \\
\theta_{0-9} & \sim \mathrm{Normal}(1, 0.1) & [\mbox{Group-level coefficient variability}]
\end{aligned}
$$
:::

@@

## R code (`brms`)

::: {.panel-tabset}

@@

:::

## Stan code

Stan code generated by `brms::stancode`.

::: {.panel-tabset}

@@

:::

# Results

@@

```{r}
#| label: results-raw
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
df %>%
    mutate(
        lv = cut_interval(lv, 5),
        age = floor(age)
    ) %>% 
    count(lv, age, response) %>%
    mutate(prop = n/sum(.$n)) %>% 
    ggplot(aes(age, prop, colour = response, fill = response)) +
    facet_wrap(~lv, nrow = 1) +
    geom_col(position = position_fill()) +
    geom_hline(yintercept = 0.5, colour = "black", size = 0.5) +
    labs(
        x = "Age (months)",
        y = "% responses", 
        colour = "Response",
        fill = "Response"
    ) +
    scale_fill_d3() +
    scale_colour_d3() +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
    theme(
        legend.position = "top",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
    )
```

@@

```{r}
#| label: results-loos
#| echo: true
#| message: false
#| warning: false
#| eval: false
#| fig-width: 7
model_loos %>% 
    as_tibble() %>% 
    rownames_to_column("model") %>% 
    relocate(
        model, 
        matches("elpd_loo"),
        matches("p_loo"), 
        matches("looic"), 
        matches("diff")
    ) %>%
    mutate_all(as.numeric) %>% 
    gt() %>% 
    tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
    tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
    tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
    tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
    fmt_number(3:13) %>% 
    cols_label(
        model = "Model",
        # elpd
        elpd_loo = md("*ELPD*"),
        se_elpd_loo = md("*SE*"),
        subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
        # p
        p_loo = md("*p*"),
        se_p_loo = md("SE"),
        subsampling_se_p_loo = md("SE<sub>sub</sub>"),
        # looic
        looic = md("*LOO-IC*"),
        se_looic = md("*SE*"),
        subsampling_se_looic = md("SE<sub>sub</sub>"),
        # diff
        elpd_diff = md("*diff*"),
        se_diff = md("*SE*"),
        subsampling_se_diff = md("*SE<sub>sub</sub>*"),
    ) 
```

@@

## Random effects

@@

@@

```{r}
#| label: results-marginal
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
nd <- expand.grid(
    n_phon_std = 0,
    freq_std = 0,
    age_std = seq(-4, 4, length.out = 30),
    doe_std = range(df$doe_std),
    lv_std = c(range(df$lv_std), 0)
)

@@
